---
title: 'Task 2: Parameter estimation'
author: "Kirsten Hodgson"
date: "3/6/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(stats)
library(broom)
library(nlstools)
library(gt)
```

### Data wrangling and exploration

```{r}
fish_catch <- read_csv(here("data", "fish_catch.csv")) 

fish_clean <- fish_catch %>% 
  rename("year" = "World Wild Fish Catch and Farmed Fish Production, 1950-2012",
         "wild_catch_mil_tons" = "X2",
         "farmed_mil_tons" = "X3",
         "total_prod_mil_tons" = "X4") %>% 
  slice(-1:-4, -68:-71) %>% 
  mutate(year_adj = 0:(n() - 1)) %>% 
  mutate(wild_catch_mil_tons = as.numeric(wild_catch_mil_tons))

#Exploratory visualization:
ggplot(data = fish_clean,
       aes(x = year, y = wild_catch_mil_tons)) +
  geom_point() +
  scale_y_continuous(limits = c(15,95),
                     breaks = seq(from = 15, to = 95, by = 5))

#Estimating *r*:
fish_exp <- fish_clean %>% 
  filter(year_adj < 35) %>% 
  mutate(ln_catch = log(wild_catch_mil_tons))

lm_r <- lm(ln_catch ~ year_adj, data = fish_exp)
#lm_r
```

The trend is described by a logistic growth relationship, with an equation like:
$P(t)=\frac{K}{1+Ae^{-rt}}$, where

- $K$ is the carrying capacity of the fishery
- $A$ is $\frac{K-P_0}{P_0}$
- $r$ is the growth rate constant

My initial estimates for $K$, $A$, and $r$ are:

- $K$ ~ $95$ (million tons)
- $A$ = $\frac{95 - 10}{10}$ ~ $8.5$
- $r$ ~ $0.03686$


### Using nonlinear least squares to estimate parameters

```{r}
wild_fish_nls <- nls(wild_catch_mil_tons ~ K/(1 + A*exp(-r*year_adj)),
              data = fish_clean,
              start = list(K = 95, A = 8.5, r = 0.03686))

#See the model summary (null hypothesis: parameter value = 0)
#summary(wild_fish_nls)

#Use broom:: functions to get model outputs in tidier format:
fish_model_out <- broom::tidy(wild_fish_nls) #Use View(model_out) to see the tidy model output.
```

After running nonlinear least squares, the estimated parameters are:

- $K$ ~ `r round(fish_model_out$estimate[1],2)` million tons
- $A$ ~ `r round(fish_model_out$estimate[2],2)` million tons
- $r$ ~ `r round(fish_model_out$estimate[3],4)` million tons/year


### Visualizing data with estimated parameters

```{r}
#Make predictions for the catch at all years in the original fish data:
wild_catch_predict <- predict(wild_fish_nls)

#Bind predictions to original data frame:
catch_complete <- data.frame(fish_clean, wild_catch_predict)

#Plot them all together:
ggplot(data = catch_complete, aes(x = year, y = wild_catch_predict, group =1)) +
  geom_line(color = "royalblue") +
  geom_point(aes(x = year, y = wild_catch_mil_tons)) +
  theme_minimal() +
  scale_y_continuous(limits = c(10,100),
                     breaks = seq(from = 10, to = 100, by = 10))  +
  labs(x = "\nYear",
       y = "Wild fish catch (millions of tons)\n",
       title = "Predicted and actual wild fish catch, 1950-2012") +
  scale_x_discrete(breaks = c(1950,1955,1960,1965,1970,1975,1980,1985,1990,1995,2000,2005,2010))
```

